% AER E 361 Mission Report Template
% Spring 2023
% Template created by Matthew Mehrtens Spring 2023

% Document Configuration DO NOT CHANGE
\documentclass[12 pt]{article}
% --------------------LaTeX Packages---------------------------------
% The following are packages that are used in this report.
% DO NOT CHANGE ANY OF THE FOLLOWING OR YOUR REPORT WILL NOT COMPILE
% -------------------------------------------------------------------

\usepackage{hyperref}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage{titling}
\usepackage{graphicx}
\usepackage{graphviz}
\usepackage[T1]{fontenc}
\usepackage{titlesec, blindtext, color} %for LessIsMore style
\usepackage{tcolorbox} %for references box
\usepackage[hmargin=1in,vmargin=1in]{geometry} % use 1 inch margins
\usepackage{float}
\usepackage{tikz}
\usepackage{svg} % Allows for SVG Vector graphics
\usepackage{textcomp, gensymb} %for degree symbol
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	urlcolor=cyan,
}
\usepackage{biblatex}
\addbibresource{lab-report-bib.bib}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{multicol}
\usepackage{array}

\usepackage{hologo} %KYR: for \BibTeX
%\usepackage{algpseudocode}
%\usepackage{algorithm}
% This configures items for code listings in the document
\usepackage{xcolor}

\usepackage{fancyhdr} % Headers/Footers
\usepackage{siunitx} % SI units
\usepackage{csquotes} % Display Quote
\usepackage{microtype} % Better line breaks

\definecolor{commentsColor}{rgb}{0.497495, 0.497587, 0.497464}
\definecolor{keywordsColor}{rgb}{0.000000, 0.000000, 0.635294}
\definecolor{stringColor}{rgb}{0.558215, 0.000000, 0.135316}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=L,
  xleftmargin=\parindent,
  language=C,
  showstringspaces=false,
  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
 }

 \lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{commentsColor}\textit,    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=tb,	                   	   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{keywordsColor}\bfseries,       % keyword style
  language=Python,                 % the language of the code (can be overrided per snippet)
  otherkeywords={*,...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=8pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{commentsColor}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{stringColor}, % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname,                  % show the filename of files included with \lstinputlisting; also try caption instead of title
  columns=fixed                    % Using fixed column width (for e.g. nice alignment)
}

\lstdefinestyle{customasm}{
  belowcaptionskip=1\baselineskip,
  frame=L,
  xleftmargin=\parindent,
  language=[x86masm]Assembler,
  basicstyle=\footnotesize\ttfamily,
  commentstyle=\itshape\color{purple!40!black},
}

\lstset{escapechar=@,style=customc}

\titlelabel{\thetitle.\quad}

% From here on out you can start editing your document
\newcommand{\subtitle}[1]{%
  \posttitle{%
    \par\end{center}
    \begin{center}\LARGE#1\end{center}
    \vskip0.5em}%
}

\newcommand{\etal}{\textit{et al}., }
\newcommand{\ie}{\textit{i}.\textit{e}., }
\newcommand{\eg}{\textit{e}.\textit{g}., }

% Define the headers and footers
\setlength{\headheight}{70.63135pt}
\geometry{head=70.63135pt, includehead=true, includefoot=true}
\pagestyle{fancy}
\fancyhead{}\fancyfoot{} % clears the headers/footers
\fancyhead[L]{\textbf{AER E 322}}
\fancyhead[C]{\textbf{Aerospace Structures Pre-Laboratory}\\
			  \textbf{Lab 6 Composite Laminate Design}\\
			  Section 4 Group 2\\
			  Matthew Mehrtens\\
			  \today}
\fancyhead[R]{\textbf{Spring 2023}}
\fancyfoot[C]{\thepage}

\begin{document}
\section*{Question 1} \label{question_1}
\textit{(\num{20} pts) Provide a concise account of composite materials: what they are in general and what type of composites are commonly used in aerostructures.}

In general, \textit{composite materials} or composites are materials composed of at least two different materials. These materials are useful because they gain the material characteristics of all the constituent materials. In aerospace engineering and aerostructures, we deal with one type of composite material in particular: fiber-reinforced composites.

Fiber-reinforced composites consist of two materials, a primary material known as the \textit{matrix} or the \textit{bonder} and a secondary material called the \textit{fiber}. The fibers are woven into the matrix in uniform patterns to create a very small layer known as a \textit{ply}. These plies are stacked on top of one another to create a composite material. Many factors can change the material properties of the composite, \eg how thick is a ply, how are the plies oriented, is the material symmetrical, which direction is the load being applied?

\section*{Question 2} \label{question_2}
\textit{(\num{3} pts) What does the subscript `s' mean in the laminate code $[0,(45,-45)_2]_s$? (\num{5} pts) How many plies in total this code indicates?}

The subscript `s' signifies the material is symmetrical. Whatever ply pattern is described within the $[]$ brackets is mirrored across the mid-plane.

The written-out form of $[0,(45,-45)_2]_s$ is \text{0/45/-45/45/-45/-45/45/-45/45/0}. When written-out, it is clear that are \num{10} plies total. The number of plies in a symmetrical composite will always be even.

\section*{Question 3} \label{question_3}
\textit{(\num{3} pts) For the fiber-reinforced composite focused in this lab, what type of material property they possess? (\num{4} pts) How many independent elastic constants are required to describe this material property?}

Fiber-reinforced composites are transversely isotropic, meaning the material properties are symmetric in two of the three directions. To describe a fiber-reinforced composite, you need four independent elastic constants, \ie the longitudinal modulus, $E_1$, the transverse modulus, $E_2$, the principal Poisson's ratio, $\nu_{12}$ or $\nu_{21}$, and the shear modulus, $G_{12}$.

\section*{Question 4} \label{question_4}
\textit{(\num{5} pts) In the theory learned in lecture notes, what kind of loads are considered/susceptible?}

In the theory, there are three types of loads considered. The resultant normal forces, $N_x$ and $N_y$; the resultant shear forces, $N_{yx}$ and $N_{xy}$; and the resultant bending moments $M_{yx}$ and $M_{xy}$.

\section*{Question 5} \label{question_5}
\textit{(\num{10} pts) What are the objectives of this lab? (\num{5} pts) What is the single strain component of particular interest? (\num{15} pts) What steps are involved in the design process?}

The objectives of this lab are to get an introduction to laminate design, determine what elements of composite design contribute most to the material properties, and to use the theory in the lecture notes in combination with the online calculator to maximize the twist curvature, $\kappa_{xy}$.
\end{document}
